\documentclass[11pt]{article}

% Change "review" to "final" to generate the final (camera-ready) version.
\usepackage[review]{acl}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}

\usepackage{inconsolata}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{pgfplots}
\pgfplotsset{compat=1.17}
\usepgfplotslibrary{groupplots}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage[most]{tcolorbox}
\usepackage{multirow}
\usepackage{url}
\usepackage{enumitem}
\usetikzlibrary{arrows.meta, positioning, calc, fit, backgrounds}

% ---------------------------------------------------------------------
% TITAN figure palette. Fixed roles, reused identically in every figure:
%   titanDSL / titanSPARQL  -- formalism identity (categorical, color-only)
%   CoT vs. NoCoT is NEVER a third color; encode it via marker/line style
%   so every figure stays legible in grayscale printing.
% ---------------------------------------------------------------------
\definecolor{titanDSL}{HTML}{2A78D6}
\definecolor{titanSPARQL}{HTML}{EB6834}
\definecolor{titanNeutral}{HTML}{52514E}
\definecolor{titanNeutralLight}{HTML}{D9D8D3}
% Ontology-figure layer colors (Section 3 only) -- deliberately NOT blue/orange,
% which are reserved for the DSL/SPARQL formalism identity everywhere else.
\definecolor{titanAdversarial}{HTML}{4A3AA7}
\definecolor{titanDefense}{HTML}{1BAF7A}

\title{TITAN: [WORKING TITLE --- to finalize] \\
  In-Context Grammars vs.\ Pretrained Query Languages \\
  for LLM-Based Multi-Hop Reasoning over Knowledge Graphs}

\author{Anonymous EACL 2027 submission}

\begin{document}
\maketitle

\begin{abstract}
TODO --- abstract is drafted last, once Sections~\ref{sec:benchmark}--\ref{sec:results} are locked, since it needs to state the four-part contribution precisely (dataset; DSL $>$ SPARQL prompted; SPARQL $>$ DSL fine-tuned; SPARQL's fine-tuned advantage is conditional, not unconditional).
\end{abstract}

\section{Introduction}
\label{sec:intro}
TODO.

\section{Related Work}
\label{sec:related}
TODO.

\input{sections/benchmark}

\input{sections/methodology}

\section{Experimental Setup}
\label{sec:expsetup}
TODO --- models, decoding, compute budget (largely a restatement/tabulation of what Section~\ref{sec:methodology} already specifies, plus anything execution-specific).

\section{Results}
\label{sec:results}
TODO --- main results table; the 2$\times$2 reversal schematic; length/operator heatmap; CoT fine-tuning reversal (diverging bars); paraphrase-robustness dumbbell chart.

\section{Discussion}
\label{sec:discussion}
TODO --- mechanism discussion for why prompting favors the DSL and fine-tuning favors SPARQL; the CoT-fine-tuning asymmetry; EM/F1 dissociation.

\section{Limitations}
\label{sec:limitations}
TODO --- RAG baseline and human/analyst plausibility evaluation are explicitly scoped out here as future work (code/infrastructure exists but was not run for this paper), not silently omitted. Also: single domain (CTI), synthetic template origin, paraphrase degradation.

\section{Conclusion}
\label{sec:conclusion}
TODO.

% No \cite commands are used in this draft; every citation point is marked
% with the literal placeholder CITAZIONE per drafting instructions.
% \bibliography{bibliography}
% \bibliographystyle{acl_natbib}

\appendix
\section{Additional Experimental Details}
\label{sec:appendix}
TODO --- per-model decoding/quantization/compute configuration; instruction-wording ablation table (Section~\ref{sec:methodology:regimes}); full length$\times$operator numeric tables; per-section dataset breakdown; full ontology node/edge/relation counts (Table~\ref{tab:ontology_counts}, referenced from Section~\ref{sec:benchmark:ontology}).

\end{document}
